The goal of this project was to build a travel mode choice model capable of embedding game theory, the simulation model can be used in further research as a starting point. The model assumes that the travelers use mixed strategy when making choices, hence, the probability of a mode being chosen is dynamic. \\

\paragraph{}
The simulations used in this work modeling a set of travelers in making choices of two modes : car or public transport. This model is based on a split mode framework often used to describe the choice behavior of travelers.
\paragraph{}
Game theory has proven its usefulness in modeling the relationship between travel mode choices and their payoffs through Nash Equilibrium. An increase in a mode's payoff can increase its proportion. The evolution part of this study is the change in Nash Equilibrium of travel mode choice.
\paragraph{}
Although people often use multimodal transport, combining two or more travel modes in the same trip, which this model lacks. This mechanism is often used to avoid traffic or lack of coverage in some areas.
\paragraph{}Finally, borrowing concepts from machine learning and artificial intelligence studies, we can use neural networks and genetic algorithms to represent the decision making process and consequent behaviors of travelers. Instead of fixed rules governing the way an individual makes decisions. It would also be interesting to add more components besides the ones that this travel mode choice model have, which can be done by further research. Studying the dynamics of travel mode choice behavior could also extract new patterns that can be out of the scope of this model.